the rethinking way of doing the binary outcome modeling

### 2021-03-16 scripts/43_comparisons_3.Rmd
finally did some real coding after 5 PM
good I tried the ordered categorical outcome not sure if i did it right tho

```{r}
pheno <- read_csv('../data/shotgun_pheno/final_comprehensive.csv') %>% 
  select(fid, center, cr_d100, toxicity, icans, crs) %>% 
  filter(center != 'I') %>% # remove the Italy ones
  gather('pheno', 'value', cr_d100:crs)

>>>>

```{r}
cur <- read_csv('../data/shotgun_pheno/final_comprehensive.csv') %>% 
  select(fid, center, cr_d100, toxicity, icans, crs) %>% 
  filter(center != 'I') %>% # manually update the label for the "not_assessed"
  mutate(cr_d100 = case_when(
    fid == '2460A' ~ 'no',
    fid == '2293A' ~ 'no',
    fid == '2563A' ~ 'no',
    TRUE ~ cr_d100
  ))

pheno <- cur %>% 
  gather('pheno', 'value', cr_d100:crs)

### 2021-03-16 scripts/54_rethinking_binary.Rmd
good I tried the ordered categorical outcome not sure if i did it right tho

```{r}
meta <- read_csv('../data/amplicon/stool/combined_2_meta.csv') %>% 
  mutate(logdiv_s = scale(log(simpson_reciprocal))) %>%  
  mutate(tox = if_else(toxicity == 'yes', 1, 0),
         cr100 = if_else(cr_d100 == 'yes', 1, 0),
         loca = if_else(center == 'M', 1, 2)) %>%  # MSK 1 ; Upenn 2
  mutate(center = factor(center),
         toxicity = factor(toxicity, levels = c('no','yes')),
         cr_d100 = factor(cr_d100, levels = c('no','yes')))

@@

```{r}
# visualize the var distribution 
table(meta$cr_d100)  
table(meta$toxicity)
#hist(meta$simpson_reciprocal)
#hist(log(meta$simpson_reciprocal)) # I think I should use log transformed
table(meta$center)
table(meta$toxicity, meta$center)
table(meta$cr_d100, meta$center)   
range(meta$logdiv_s)
```

```{r}
set.seed(123)
dat_list <- list(
    tox = meta$tox,
    cr100 = meta$cr100,
    location = meta$loca,
    logdiv_s = meta$logdiv_s
    )

@@

@@

```{r}
# the raw proportion of yes toxicity in each place
rawtox <- meta  %>% 
  count(center, toxicity) %>% 
  left_join(meta %>% 
              count(center, name = 'total'), by = 'center') %>% 
  group_by(center, toxicity) %>% 
  summarise(frac = round(n/total, 2)) %>% 
  filter(toxicity == 'yes')

@@

bind_rows(
  tibble(value = pl, grp = 'lowest'),
  tibble(value = ph, grp = 'highest')
) %>% 
  gghistogram(x = 'value',bins = 30, fill = 'grp', palette = 'nejm', color = 'white', add = 'mean', xlab = 'Predicted probability of CR d100', ylab = 'Probability density',
              title = 'Predicted probability of CR d100 between the pt with highest div and the pt with lowest') 
```

>>>>

```{r}
meta <- read_csv('../data/amplicon/stool/combined_2_meta.csv') %>% 
  mutate(logdiv_s = scale(log(simpson_reciprocal))) %>%  
  mutate(tox = if_else(toxicity == 'yes', 1, 0),
         cr100 = if_else(cr_d100 == 'yes', 1, 0),
         loca = if_else(center == 'M', 1, 2)) %>%  # MSK 1 ; Upenn 2
  mutate(center = factor(center),
         toxicity = factor(toxicity, levels = c('no','yes')),
         cr_d100 = factor(cr_d100, levels = c('no','yes'))) %>%
  # have a 2 cutpoint for the CRS
  mutate(crs3 = case_when(
    crsgrade == 0 ~ 1,
    crsgrade == 1 ~ 2,
    crsgrade >= 2 ~ 3
  ))

@@

```{r}
# visualize the var distribution 
table(meta$cr_d100)  
table(meta$toxicity)
#hist(meta$simpson_reciprocal)
#hist(log(meta$simpson_reciprocal)) # I think I should use log transformed
table(meta$center)
table(meta$toxicity, meta$center)
table(meta$cr_d100, meta$center)   
range(meta$logdiv_s)
table(meta$crs3)
table(meta$crsgrade, meta$center) 
```

```{r}
set.seed(123)
dat_list <- list(
    tox = meta$tox,
    crs3 = meta$crs3,
    cr100 = meta$cr100,
    location = meta$loca,
    logdiv_s = meta$logdiv_s
    )

@@

```{r}
# another model with the crs grade 0,1,>=2 
mcrs3 <- ulam(
    alist(
        crs3 ~ dordlogit( phi , cutpoints ),
        phi <- b*logdiv_s + a[location] ,
        b ~ dnorm( 0 , 1),  
        a[location] ~ dnorm( 0 , 0.5 ),
        cutpoints ~ dnorm( 0 , 1.5 )
    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
plot(precis(mcrs3, depth = 2, prob = 0.95) ) 
# the cutpoints are not of much interet, this suggests that the diversity really isn't having a correlation?
```

@@

```{r}
# the raw proportion of yes toxicity in each place
# rawtox <- meta  %>% 
#   count(center, toxicity) %>% 
#   left_join(meta %>% 
#               count(center, name = 'total'), by = 'center') %>% 
#   group_by(center, toxicity) %>% 
#   summarise(frac = round(n/total, 2)) %>% 
#   filter(toxicity == 'yes')

@@

bind_rows(
  tibble(value = pl, grp = 'lowest'),
  tibble(value = ph, grp = 'highest')
) %>% 
  gghistogram(x = 'value',bins = 30, fill = 'grp', palette = 'nejm', color = 'white', add = 'mean', xlab = 'Predicted probability of CR d100', ylab = 'Probability density',
              title = 'Predicted probability of CR d100 between the pt with highest div and the pt with lowest')

mean(pl)
mean(ph)
```
